---
title: "Week7 - Practice Quiz"
author: "Ashutosh Agarwal"
date: "7/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practice Quiz

# Question 1, 2,3 :

Use this model to obtain an estimate of the change in mean tooth length for a dose increase of 1 milligram per day, when the supplement type is orange juice.

```{r}
model = lm(len ~ dose * supp, data = ToothGrowth)

slopeOrangeJuice = model$coefficients["dose"]
interceptOrangeJuice = model$coefficients["(Intercept)"]

slopeVC = model$coefficients["dose"] + model$coefficients["dose:suppVC"]
interceptVC = model$coefficients["suppVC"]
```

- **Question 1:**

```{r}
slopeOrangeJuice
```

- **Question 2:**

```{r}
slopeVC
```

- **Question 3:**

To test for interaction between does and supplier type, we need to remove interaction term (with coefficient $\beta_3$) to get null model.

Since both null model and full model differ by only one coefficient, we can use t-test on $\beta_3$

```{r}
summary(model)$coefficients["dose:suppVC", "Pr(>|t|)"]
```

# Question 4

```{r}
unique(ToothGrowth$dose)

# Add a new column that contains factored representation of dose variable.
ToothGrowth$doseFactor = as.factor(ToothGrowth$dose)

# Now train a model
# We are using plus (+) instead of asterisk (*) because we are not interested in interaction.
model = lm(len ~ doseFactor + supp, data = ToothGrowth)

oneMgVC = model$coefficients["(Intercept)"] + model$coefficients["doseFactor1"] + model$coefficients["suppVC"]
twoMgVC = model$coefficients["(Intercept)"] + model$coefficients["doseFactor2"] + model$coefficients["suppVC"]
oneMgOJ = model$coefficients["(Intercept)"] + model$coefficients["doseFactor1"]
twoMgOJ = model$coefficients["(Intercept)"] + model$coefficients["doseFactor2"]

oneMgVC - twoMgVC
oneMgOJ - twoMgOJ
```

# Question 5 - Parameterization.

```{r}
ToothGrowth$d1 = as.numeric(ToothGrowth$dose == 0.5)
ToothGrowth$d2 = as.numeric(ToothGrowth$dose == 1)
ToothGrowth$d3 = as.numeric(ToothGrowth$dose == 2)
ToothGrowth$svc = as.numeric(ToothGrowth$supp == "VC")

model = lm(len ~ 0 + d1 + d2 + d3 + svc, data = ToothGrowth)
model$coefficients["d3"]
```

