---
title: "Final Project Proposal"
author: "STAT 420, Summer 2021, Ashutosh Agarwal, Chandan Goel"
date: '07/20/2021'
output:
  html_document: 
    theme: readable
    toc: yes
  pdf_document: default
urlcolor: cyan
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Team

  - **Ashutosh Agarwal (aa61)**
  - **Chandan Goel (chandan3)**

## Project Title

**House Price Prediction in King County, WA**

## Background information of data set

TBD by Chandan

## Brief statement

TBD by Chandan

## Data file description

  * **Source:** This data set is taken from [Kaggle](https://www.kaggle.com/harlfoxem/housesalesprediction).
  
  * **Brief description:** This data set contains house prices for King County, WA that were sold between `May 2014` and `May 2015`
  
  * **Loading data file**
  
```{r}
data = read.csv("kc_house_data.csv")
head(data)
```
  
  - **Summary of data**
  
  This data set contains **`r nrow(data)` observations**. Each observation has **`r ncol(data)` variables**.
  
  Following are some variables that we think are of great importance and we will most likely include in our model.
  
```{r}
metadata = read.csv("Metadata.csv")
# rownames(metadata) = NULL
knitr::kable(metadata[metadata$IsImportant == "Yes", -5], row.names = FALSE)
```